You are the way you structurally talk: structural-temporal neighbourhoods of posts to characterize users in online forums

### 2016-04-13 doc/IEEEtran/paper_IEEEformat.tex
paper ++

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
The abstract goes here.
\end{abstract}

@@

\section{Introduction}
The popularization of online forums has brought a growing interest on their underlying dynamics. As any other complex system, the dynamic of online forums can be studied at different levels, from the most macro to the most micro. Macro dynamics are, for instance, the evolution of some global properties of the social graph such as its diameter, or its distribution degree. Micro dynamics are, for instance, the triadic motifs that represent local phenomena such as transitivity (friends of my friends are also my friends).

An interesting question in online communities is that concerning roles. In sociology, roles are generally seen as the set of expected behaviours that are attached  to a position in the community. Extrapolating the notion of role, some researchers have looked for roles in online forums. Some others have tried to detect the roles and the users who hold that roles.

@@

@@

The remaining of the paper is as follows. We first discuss about the convenience of the classic neighbourhood definition in dynamic graphs such as discussion trees. Then we introduce two new definitions that takes into account the time to overcome some of the limitations of the structural neighbourhood. We will finally apply these time-based neighbourhoods to detect clusters of users than tend to appear in the same time-based neighbourhoods.

\begin{figure}
	\centering
	%\includegraphics[width=0.3\textwidth]{tree1}
	\includegraphics[width=0.3\textwidth]{tree2}
	%\includegraphics[width=0.3\textwidth]{tree3}
	\caption{Representation of discussion threads with post tree graphs. Vertex represent posts and edges represent replies between posts.}
	\label{fig:trees}
\end{figure}

\section{Discussion trees}
We represent a conversation thread as a tree graph where vertex represent posts and edges represent replies from some post to another. The tree is rooted at the post who started the thread. Figure~\ref{fig:trees} shows some real examples of trees in a Reddit\footnote{www.reddit.com} forum.

@@

\section{Structural neighbourhoods in discussion trees}
Extending the classic definition of neighbourhood according to which two vertices are neighbours if the distance between them is one, in this paper we define the \textit{structural neighbourhood} as follows:

\begin{definition}
Given a tree graph $G$, the \textit{structural neighbourhood} of radius $r$ of post $i$, denoted as $\mathcal{N}_i(r)$, is the induced graph composed of all the vertices that are at distance equal or less than $d$ from post $i$.
\end{definition}
% pros and cons
This definition has two drawbacks when used in the context of conversation trees.
%%% not all dynamics are captured
First, the dynamics of the conversation (time or order in which posts are attached to the tree) are not entirely captured in the structure of a tree representation. We know, for instance, that the time in which a node was attached to the tree is always posterior to that of its parent. But it impossible to say, by just looking at the structure of the tree, the order in which a set of sibling posts replied to their common parent post. Thus, a single \textit{structural neighbourhood} may sometimes correspond to very different dynamics.
%TODO: ejemplo artificial para demostrar que la definicion de vecindario estructural puede corresponder a dinamicas muy diferentes.
%% unbounded size
Second, a structural neighbourhood at a given radius $r$ has an unbounded number of posts, and therefore the space of possible neighbourhoods is infinite. This poses a problem when trying to categorize conversations since many conversations, while structurally different, can be considered semantically equivalent (see Figure \ref{fig:large_neighbourhood}).

\begin{figure}
	\centering
	\includegraphics[width=0.24\textwidth]{large_neighbourhood}
	\includegraphics[width=0.24\textwidth]{small_neighbourhood}
	\caption{The size of the neighbourhood with radius $r$ is unbounded. These two graph represent frequent neighbourhoods of a post (red) that replied to the root (white). However, the tree in the left corresponds to a very successful root while the three in the right has not brought the attention of too many users.}
	\label{fig:large_neighbourhood}
\end{figure}

\section{Structuro-temporal neighbourhoods in discussion trees}
In this section, we propose two new time-based definition of neighbourhoods.

\subsection{Order-based}
Our first definition is based on the order in which posts are attached to the tree. 
\begin{definition}
Given a tree graph $G$, the \textit{structural-temporal neighbourhood} of distance $d$ of post $i$, denoted as $\mathcal{N}_{i}^T(r,n)$ is the induced subgraph from its structural neighbourhood composed the $n$ vertices that are closer to $i$ in time and for which there exists a path to $i$ in $\mathcal{N}_{i}^T(r,n)$.  
\end{definition}

This definition has two advantages over the \textit{structural neighbourhood}. First, the temporal aspect of the conversation is better taken into account since the neighbourhood only includes posts that are structurally and temporally close. Second, the size of the neighbourhood has an upper bound of $max(\mathcal{N}_i(r)), o)$.

\subsection{Time-based}
Our second definition uses the real timestamps of posts in order to decide where the boundaries of a neighbourhood are. We might naively set fixed time-based boundaries for the neighbourhood by only including in the neighbourhood those posts whose timestamp $p_i$ is at distance less than $\tau$ from the ego post $|t_i-t_{ego}|<\tau$. However, the pace at which posts are added to the conversation may be very different between conversations, and even between different parts of the same conversation tree. Instead, we propose a definition based on the concept of $\textit{local dynamic}$.

\begin{definition}
The temporal neighbourhood $\mathcal{N}_{i}^T(r)$ is the maximal subgraph of the structural neighbourhood $\mathcal{N}_i(r)$ where all the posts belong to the same local dynamic than the ego post.
\end{definition}
Our definition of \textit{local dynamic} is based on the detection of changepoints in the time stamps of the posts. Given a sequence of consecutive posts in the same branch $b$ of a tree, denoted as $p_s,...,p_e$ we say that they belong to the same vertical dynamic if there is no (vertical) changepoint $p_i$ in $b$ such that $p_1 \prec p_i \preceq p_n$. Similarly, given a sequence of chronologically sorted  siblings $s$, denoted as $p_s,...,p_e$, we say that they belong to the same horizontal dynamic if there is no (horizontal) changepoint $p_i$ in $s$ such that $p_1 \prec p_i \preceq p_n$. Now we can outline the complete algorithm for time-based neighbourhood (Algorithm \ref{alg:temporal_neighbourhood}). Figure \ref{fig:cutpoints} illustrate a tree with a set of breakpoints and the temporal neighbourhood for a given node.
\begin{algorithm}[H]
\begin{algorithmic}
\REQUIRE Posts tree $g$, vertical breakpoints, horizontal breakpoints, ego post $ego$
\ENSURE Subgraph of $g$ with all vertices in $V(g)$
\STATE Compute structural neighbourhood $\mathcal{N}_i(r)$
\STATE ancestors $\leftarrow$  ancestors(ego) in $\mathcal{N}_i(r)$
\STATE older\_siblings $\leftarrow$ older\_siblings(ego) in $\mathcal{N}_i(r)$
\STATE dump $\leftarrow \varnothing$
\FOR{bp $\in$ vertical breakpoints}
 \IF{bp $\in$ ancestors}
   \STATE dump $\leftarrow$ dump $\cup$ ancestors(bp)
 \ELSE
   \STATE dump $\leftarrow$ dump $\cup$ descendants(bp) $\cup$ bp
  \ENDIF
\ENDFOR
\FOR{bp $\in$ horizontal breakpoints}
  \IF{bp $\in$ (older\_siblings $\cup$ ancestors)}
     \STATE dump $\leftarrow$ dump $\cup$ older\_siblings(bp)
   \ELSE
     \STATE dump $\leftarrow$ dump $\cup$ younger\_siblings(bp) $\cup$ bp
  \ENDIF
\ENDFOR
\STATE $\mathcal{N}_i^{(t)}(r) \leftarrow$ delete(dump\_posts) from $\mathcal{N}_i(r)$
\end{algorithmic}
\caption{Extraction of time-based neighbourhood}
\label{alg:temporal_neighbourhood}
\end{algorithm}

\begin{figure}
\centering
\includegraphics[width=0.45\textwidth]{breakpoints}
\caption{Time-based neighbourhood. Horizontal changepoints (yellow) and vertical changepoints (orange) represent posts that are temporally far from their predecessors (siblings or parents) and therefore set the limits of the neighbourhood. Post 6 is the ego, and posts with circular borders are those belonging to the time-based neighbourhood.}
\label{fig:cutpoints}
\end{figure}

\subsection{Colored neighbourhoods}
In conversation trees, the root vertex plays an important role since it represents the post that opened the conversation and thus set the topic of the thread. Therefore, a reply to the root post is clearly different from a reply to any other post in the tree. We distinguish root posts from non-root posts by assigning root posts a unique color (white). We also assign a unique color to the ego post (red).

\subsection{Neighbourhood pruning}\label{sec:pruning}

@@

\begin{figure*}
	\centering
	\includegraphics[width=0.8\textwidth]{census_orderbased_1}
	%\includegraphics[width=0.8\textwidth]{census_orderbased_2}
	%\includegraphics[width=0.8\textwidth]{census_orderbased_3}
	\caption{Most frequent order-based neighbourhoods with $r=2$ and $n=4$}
	\label{fig:census_orderbased}
\end{figure*}

@@

\begin{figure*}
	\centering
	\includegraphics[width=0.8\textwidth]{neighbourhoods_time_1}
	\includegraphics[width=0.8\textwidth]{neighbourhoods_time_2}
	%\includegraphics[width=0.8\textwidth]{neighbourhoods_time_3}
	%\caption{Time-based neighbourhoods (most frequent)}
	\label{fig:neighbourhoods_time}
\end{figure*}

@@

@@

@@

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
\begin{thebibliography}{1}

\bibitem{IEEEhowto:kopka}
H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.

\end{thebibliography}

>>>>

% use for special paper notices
\IEEEspecialpapernotice{(Draft version 13/04/2016)}

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}Users of social networks are often characterised by extracting some relevant features from the graphs associated to that network. The structural dynamic of the conversation are usually forgotten due to a lack of proper tools. We present a purely graph-based method to cluster users in online forums. 
\end{abstract}

@@

\section{Introduction}
The popularization of online forums has brought a growing interest on their underlying dynamics. As any other complex system, the dynamic of online forums can be studied at different levels, from the most macro to the most micro. Macro dynamics are, for instance, the evolution of some global properties of the social graph such as its diameter, or its distribution degree. Micro dynamics are, for instance, the triadic motifs that represent local phenomena such as transitivity (friends of my friends are also my friends) and that may explain some macro phenomena.

The study of user behaviours has also brought the attention of researchers. 
An interesting question in online communities is that concerning roles. In sociology, roles are generally seen as the set of expected behaviours that are attached  to a position in the community. Extrapolating the notion of role, some researchers have looked for roles in online forums. Some others have tried to detect the roles and the users who hold that roles.

@@

In this paper we present two types of structural patterns adapted to discussion trees. These trees can represent e-mails, forums, etc. Here we will use forum data to illustrate out methodology.

@@

The remaining of the paper is as follows. We first discuss about the convenience of the classic neighbourhood definition in dynamic graphs such as discussion trees. Then we introduce two richer definitions that take into account the order and the time. To illustrate the kind of neighbourhoods we obtain in a real forum, we show the neighbourhood census in a Reddit forum. Finally, we apply both neighbourhood definitions to detect clusters of users that tend to appear in the same neighbourhoods.
We close the paper with some suggestions of future research.
\begin{figure}
	\centering
	%\includegraphics[width=0.3\textwidth]{tree1}
	\includegraphics[width=0.3\textwidth]{tree2}
	%\includegraphics[width=0.3\textwidth]{tree3}
	\caption{Representation of discussion threads with post tree graphs. Vertex represent posts and edges represent replies between posts.}
	\label{fig:trees}
\end{figure}

\section{Discussion trees}
The interactions like those found in an online forum can be modelled, at least, by two sort of graphs. On the one hand, we can create a graph where vertices represent individuals and an edges represent an interaction between two individuals. At the individual level, this graph is suitable for a large set of analysis concerning the centrality of users, the detection of communities, the detection of groups that tend to interact with the same other groups. At the community level we can analyse properties such as the density of the graph, the clustering coefficient or the degree distribution. On the other hand, we can create a separate tree graph for every discussion where posts are represented by the vertices and and edge from one vertex to another indicates that first post is a reply to the second. This allows to analyse things like patterns of time delay between posts \cite{Bhatt2012}.

We can even mix different levels of representations to study the relationship between the properties and dynamics among the different levels \cite{Dorat2007}.

Of course, we are not limited to the study of graphs and some studies take into account the content of the conversations, the cross-posting activity of users, and so forth \cite{Whittaker1998}

@@

Others: \cite{Gaumont2016}

\section{Structural neighbourhoods}
Extending the classic definition of neighbourhood according to which two vertices are neighbours if the distance between them is one, we start by the following definition of \textit{structural neighbourhood}:

\begin{definition}
Given a tree graph $G$, the \textit{structural neighbourhood} of radius $r$ of post $i$, denoted as $\mathcal{N}_i(r)$, is the induced graph composed of all the vertices that are at distance equal or less than $d$ from post $i$.
\end{definition}
% pros and cons
In the context of discussion threads, this definition has two limitations. First,

 the decision on whether to include some post in the neighbourhood is only based on the structural distance, and therefore two posts that are at distance $d\leq r$ are considered neighbours of $i$ regardless of the time they were written. In conversations, time plays an important role, therefore this is an important limitation. Another consequence of looking only at the structure is that the number of possible neighbourhoods within a radius $r$ is infinite. This poses a problem when trying to categorize conversations since many conversations, while structurally different, can be considered semantically equivalent.

\section{Structuro-temporal neighbourhoods}
As we have discussed, structural neighbourhood falls short in the analysis of conversational structures. Instead, we propose two new definitions of neighbourhood that take into account order and time in which posts are written.

\begin{figure*}
\centering
\includegraphics[width=0.45\textwidth]{order_neighbourhood}
\includegraphics[width=0.45\textwidth]{breakpoints}
\caption{Illustration of order-based (left) and time-based (right) neighbourhoods. Grey nodes represent the ego post. Nodes with circles are those in the neighbourhood. Numbers indicate the order of the post. The order-based neighbourhood has parameters $r=3$, $n=6$. In the time-based neighbourhood, horizontal changepoints (yellow) and vertical changepoints (orange) represent posts that are temporally far from their predecessors (siblings or parents) and therefore set the limits of the neighbourhood.}
\label{fig:cutpoints}
\end{figure*}
\subsection{Order-based}
Our first definition is based on the order in which posts are attached to the tree. 
\begin{definition}
Given a tree graph $G$, the \textit{structural-temporal neighbourhood} of distance $d$ of post $i$, denoted as $\mathcal{N}_{i}^T(r,n)$ is the induced subgraph from its structural neighbourhood composed the $n$ vertices that are closer to $i$ in time and for which there exists a path to $i$ in $\mathcal{N}_{i}^T(r,n)$.  
\end{definition}

This definition has two advantages over the \textit{structural neighbourhood}. First, the temporal aspect of the conversation is better taken into account since the neighbourhood only includes posts that are structurally and temporally close. Second, the size of the neighbourhood has an upper bound of $max(\mathcal{N}_i(r)), n)$ thus making the space of possible neighbourhood structures finite.

\subsection{Time-based}
Our second definition uses the real timestamps of posts in order to automatically decide where the boundaries of a given neighbourhood are. We might naively set fixed time-based boundaries for the neighbourhood by only including in the neighbourhood those posts whose timestamp $p_i$ is at distance less than $\tau$ from the ego post $|t_i-t_{ego}|<\tau$. However, the pace at which posts are added to the conversation may be very different between conversations, and even between different parts of the same conversation tree, making it difficult to set a fixed temporal radius $\tau$. Instead, we propose a definition based on the concept of $\textit{local dynamic}$.

Our definition of \textit{local dynamic} is based on the detection of changepoints in the time stamps of the posts (ref). Given a sequence of consecutive posts in the same branch $b$ of a tree, denoted as $p_s,...,p_e$ we say that they belong to the same vertical dynamic if there is no (vertical) changepoint $p_i$ in $b$ such that $p_1 \prec p_i \preceq p_n$. Similarly, given a sequence of chronologically sorted  siblings $s$, denoted as $p_s,...,p_e$, we say that they belong to the same horizontal dynamic if there is no (horizontal) changepoint $p_i$ in $s$ such that $p_1 \prec p_i \preceq p_n$.

\begin{definition}
Given an ordered tree $T$, the temporal neighbourhood $\mathcal{N}_{i}^T(r)$ is the maximal subgraph of the structural neighbourhood $\mathcal{N}_i(r)$ where all the posts belong to the same local dynamic than the ego post.
\end{definition}
Now we can outline the complete algorithm for time-based neighbourhood (Algorithm \ref{alg:temporal_neighbourhood}). Figure \ref{fig:cutpoints} illustrate a tree with a set of breakpoints and the temporal neighbourhood for a given node.
\begin{algorithm}[H]
\begin{algorithmic}
\REQUIRE Posts tree $g$, vertical breakpoints, horizontal breakpoints, ego post $ego$
\ENSURE Subgraph of $g$ with all vertices in $V(g)$
\STATE Compute structural neighbourhood $\mathcal{N}_i(r)$
\STATE ancestors $\leftarrow$  ancestors(ego) in $\mathcal{N}_i(r)$
\STATE older\_siblings $\leftarrow$ older\_siblings(ego) in $\mathcal{N}_i(r)$
\STATE dump $\leftarrow \varnothing$
\FOR{bp $\in$ vertical breakpoints}
 \IF{bp $\in$ ancestors}
   \STATE dump $\leftarrow$ dump $\cup$ ancestors(bp)
 \ELSE
   \STATE dump $\leftarrow$ dump $\cup$ descendants(bp) $\cup$ bp
  \ENDIF
\ENDFOR
\FOR{bp $\in$ horizontal breakpoints}
  \IF{bp $\in$ (older\_siblings $\cup$ ancestors)}
     \STATE dump $\leftarrow$ dump $\cup$ older\_siblings(bp)
   \ELSE
     \STATE dump $\leftarrow$ dump $\cup$ younger\_siblings(bp) $\cup$ bp
  \ENDIF
\ENDFOR
\STATE $\mathcal{N}_i^{(t)}(r) \leftarrow$ delete(dump\_posts) from $\mathcal{N}_i(r)$
\end{algorithmic}
\caption{Extraction of time-based neighbourhood}
\label{alg:temporal_neighbourhood}
\end{algorithm}

\subsection{Colored neighbourhoods}
Once we extract a neighbourhood, we assign colors to their nodes to represent some relevant characteristics. In particular, we assign the red color to the ego post, the orange colour to all other posts written by the author of the ego (to identify re-entries) a white color to the root (if included in the neighbourhood), the yellow color to all posts written by the same author who wrote the parent of ego (to identify, for instances, debates between the ego author and the other author), the grey color to all posts written before the ego and the black color to all posts written after the ego. Since some posts might correspond to several colors, we perform a sequential assignment of colors given by $\texttt{grey, black, orange, ego, yellow, white}$.

\begin{figure*}
\centering
\includegraphics[width=0.2\textwidth]{neighbourhood_time_342}
\includegraphics[width=0.19\textwidth]{neighbourhood_time_456}
\includegraphics[width=0.18\textwidth]{neighbourhood_time_501}
\includegraphics[width=0.18\textwidth]{neighbourhood_time_524}
\caption{Illustration of different real structures using different colors for the root (white), the ego (red) and other posts written by the same author (orange) the ego parent and other posts written by the same author (yellow), posts previous to the ego (grey) and posts posterior to the ego.}
\label{fig:cutpoints}
\end{figure*}

\subsection{Neighbourhood pruning}\label{sec:pruning}
Although the two proposed definitions limit the neighbour space considerably, we can still find structures whose only difference is that one of them has more leafs hanging from some of the nodes. Thus, we prune every neighbourhood by leaving a maximum of two consecutive siblings of the same color and no children.

@@

\begin{figure*}
	\centering
	\includegraphics[width=0.8\textwidth]{census_orderbased_1}
	\includegraphics[width=0.8\textwidth]{census_orderbased_2}
	%\includegraphics[width=0.75\textwidth]{census_orderbased_3}
	\caption{Most frequent order-based neighbourhoods with $r=2$ and $n=4$}
	\label{fig:census_orderbased}
\end{figure*}

@@

\begin{figure*}
	\centering
	\includegraphics[width=0.8\textwidth]{census_timebased_1}
	\includegraphics[width=0.8\textwidth]{census_timebased_2}
	%\includegraphics[width=0.8\textwidth]{neighbourhoods_time_3}
	\caption{Most frequent time-based neighbourhoods}
	\label{fig:neighbourhoods_time}
\end{figure*}

@@

\begin{figure*}
	\centering
	\includegraphics[width=0.9\textwidth]{PCA_cluster_orderbased}
	\includegraphics[width=0.9\textwidth]{PCA_cluster_timebased}
	\caption{PCA projections of the order-based (above) and time-based (below) neighbourhood features}
	\label{fig:neighbourhoods_time}
\end{figure*}

@@

% Maybe mention that we can take for instance only the posts before the ego and use it to understand when when can recommend a post based on the structure.

@@

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
\bibliography{bibliography}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
%\begin{thebibliography}{1}
